### 2018-02-04 Physics Notes/Turning Points/Electrons/Electrons.tex
Notes Update
Physics Notes Update

$$eV_a=\frac{1}{2}mv^2 \quad \frac{e}{m}=\frac{\frac{1}{2}v^2}{v_a}$$

@@

$$\frac{e}{m}=\frac{1}{2v_a}\Bigg(\frac{E}{B}\Bigg)^2$$

@@

$$\textrm{This is the centripetal force} \frac{mv^2}{R}$$

@@

>>>>

$$eV_a=\frac{1}{2}mv^2 \quad \frac{e}{m}=\frac{\frac{1}{2}v^2}{\textrm{v}_a}$$

@@

$$\frac{e}{m}=\frac{1}{2\textrm{v}_a}\Bigg(\frac{E}{B}\Bigg)^2$$

@@

$$\textrm{This is the centripetal force,} \quad \frac{mv^2}{R}$$

@@

\underline{\huge Questions from past papers}
\textit{Why does an oil droplet reach a constant speed when the plate voltage is switched off in Milikan's experiment}
Weight pulls droplet down, reaches terminal velocity
$ $\\
\textit{Derive $r^2=\frac{9\eta v}{2\rho g}$}
$ $\\
\textit{Determine the mass or radius of a droplet in Milikan's experiment}\\
$$\textrm{Speed}=\frac{\textrm{Distance}}{\textrm{Time}}$$
$$r=\sqrt{\frac{9\eta v}{2\rho g}}$$
$$m=\frac{4}{3}\pi r^3\rho$$
$ $\\
\textit{Charge from mass}
$$Q=\frac{mgd}{V}$$
$ $\\
\textit{Determine the sign of the charge carried by a stationary droplet in Milikan's experiment}
Opposite to the charge of the top plate to ensure forces are balanced
$ $\\
\textit{Comment on the significance of a charge calculated in Milikan's experiment}
It will be a multiple of the charge of an electron
\section{Electron Guns}
\textit{Speed of electrons from the voltage they are accelerated through a Voltage}\\
$$\frac{1}{2}mv^2=eV$$
$$v=\sqrt{\frac{eV}{\frac{1}{2}m}}$$
$ $\\
\textit{Why are there two power supplies in an electron gun?}
One to cause thermionic emission, one to accelerate the electrons
\section{Electrons in magnetic fields}
\textit{Does the speed of an electron change in a magnetic field}
No as force is perpendicular to velocity, meaning no work is done
$ $\\
\textit{Specific charge from radius of curvature, magnetic flux density and accelerating voltage}
$$\frac{1}{2}mv^2=eV$$
$$\frac{mv^2}{r}=BeV$$
$$\frac{e}{m}=\frac{2V}{B^2r^2}$$
$ $\\
\textit{What is the velocity of an electron when magnetic and electric fields are equal}
$$EQ=BQv$$
$$v=\frac{E}{B}$$
$ $\\
\textit{Velocity of electrons in a magnetic field}
$$\frac{mv^2}{r}=Bev$$
$$v=\frac{Ber}{m}$$
$ $\\
\textit{Why do electrons undergo circular motion in a magnetic field}
Force perpendicular to motion, does not change speed, causes direction of motion to change
$ $\\
\section{Electrons in electric fields}
\textit{Draw the path of an electron beam in an electric field}
Curves up while between plates, straight line continuing up after leaving plates
$ $\\
\textit{Specific charge of an electron from curving in an electric field}
$$ma=\frac{eV}{d}$$
Calculate acceleration using SUVAT
$$\frac{ad}{V}=\frac{e}{m}$$
$ $\\
\textit{Specific charge from electric field strength}
$$EQ=mg$$
$$\frac{Q}{m}=\frac{g}{E}$$
$ $\\
\textit{Why does an electron beam in an electric field curve at an increasing angle to the horizontal}
The electrons feel a constant force, meaning they have constant vertical acceleration, causing the vertical component of the velocity to increase, while the horizontal component is constant, causing an increasing angle.
\textit{Explain thermionic emission}
The current heats the wire, meaning the electrons in the filament gain sufficient k.e. to leave the filament
$ $\\
\textit{Why must the filament be in an evacuated tube for thermionic emission}
So that there are no collisions
$ $\\
\textit{Why is thermionic emission negligible when the filament current is too low}
If the current is too low there is little heating effect, meaning that the electrons don't gain enough kinetic energy to be released
\section{Discharge Tubes}
\textit{Why is light emitted in a Crooke's tube}
Electrons present, collide with atoms, excitation occurs, emit photon as they return to the ground state
$ $\\
\textit{Why can a glow not be seen until the pressure is low enough in a Discharge tube}
The electrons can't gain enough kinetic energy as there are too many molecules present
$ $\\
